Gringo 2.0

### 2008-09-02 doc/documentation/paper.tex
Doku: changed blocksworld example to rolands new version
git-svn-id: https://potassco.svn.sourceforge.net/svnroot/potassco/branches/gringo2@483 3e02f596-3131-0410-8352-e04eeaec37cc

 1{move(X,Y,k) : block(X) : location(Y) : X != Y}1
\end{verbatim}
A move is of the form ``\texttt{move(X,Y,k)}'' where \texttt{X} is the block that is moved, \texttt{Y} the new position of the block and \texttt{k} the step when the moved is executed.
\newline
Now that we have exactly one move per step we need to define what happens if we move a block.
Understandably if we move block \texttt{X} to position \texttt{Y} in step \texttt{k}, the new location of the block shall be \texttt{Y}.
\begin{verbatim}
 on(X,Y,k) :- move(X,Y,k), block(X), location(Y).
\end{verbatim}
For all other blocks that have not moved, the position shall be the same.
In ASP a little trick is used to describe such a condition.
We first define a little helper predicate that states that a block is not on a certain position.
\begin{verbatim}
 not_on(X,Y,k) :- on(X,Z,k), Y != Z, block(X), location(Y;Z).
\end{verbatim}
Now block \texttt{X} is ``\texttt{not\_on}'' a position, where it is not ``\texttt{on}''.
With this helping predicate we can define that a block is on the same position in step \texttt{k}, if it was not ``\texttt{not\_on}'' that position in the step before.
\begin{verbatim}
 on(X,Y,k) :- on(X,Y,k-1), block(X), location(Y), not not_on(X,Y,k).
\end{verbatim}
This trick is often used to describe the state of something that has not changed by an action.
\newline
Now we have to restrict our moves, because only one block can be moved at the same time and it can only be moved if no other block is on top of it.
To restrict the moves, an integrity constraint is used.
\begin{verbatim}
 :- move(X,Y,k), on(Z,X,k-1), block(X;Z), location(Y).
\end{verbatim}
So it is not allowed to move block \texttt{X} in step k, if there is another block on it in the state before.
Also you can not move a block to a position where another block is on, so only one block on top of another is allowed.
\begin{verbatim}
 :- move(X,Y,k), on(Z,Y,k-1), block(X;Y;Z).
\end{verbatim}
The description of the problem is nearly finished. All that has to be done is to formulate a query.
As a constraint to our solution we want to stop if all blocks are on top of each other in reversed order.
This query shall only hold in the last step, therefore we write it in the volatile part.
As always for constraints, we use an integrity constraint to express that all solutions are false where there is not block $n+1$ on the block $n$.
\begin{verbatim}
 #volatile k.

 :- not on(X+1,X,k), block(X), X < n.
\end{verbatim}
To run this example, iclingo has to be used and the constant $n$ has to be set via command line.
A potential command line call could look like
\begin{verbatim}
 iclingo --const n=10 blocksworld.lp
\end{verbatim}
and the answer to the problem can be read from the answer set.

%  1 #hide.
%  2 #show mv(X,Y,Z).
%  3
%  4 #base.
%  5
%  6 bl(1..n).
%% Es gibt "n" Blöcke
%  7 loc(t).
%  8 loc(X) :- bl(X).
%% und die Positionen t=table, 1 ... n
%  9 
%% Block 1 ist auf Block 2, 2 auf 3 ... und n auf dem Tisch
% 11
% 12 #cumulative k.
% 13
% 14 1{mv(X,Y,k) : bl(X) : loc(Y) : X != Y}1.
%% Wähle jeden Zug einen move von Block X nach Y aus
% 15 on(X,Y,k) :- mv(X,Y,k), bl(X), loc(Y).
%% Ein Block X ist auf der Position Y (im Schritt k)falls block X nach Y bewegt wird(in Schritt k).
% 16 on(X,Y,k) :- on(X,Y,k-1), bl(X), loc(Y), not non(X,Y,k).
% 17 non(X,Y,k) :- on(X,Z,k), Y != Z, bl(X), loc(Y;Z).
%% Ein Bock X ist auf Position Y (in Schritt k) wenn er nicht, nicht dort lag.
%% non(X,Y) gilt im Schritt k, falls X NICHT auf Y liegt.
% 18 :- mv(X,Y,k), on(Z,X,k-1), bl(X;Z), loc(Y).
%% move von X nach Y (im Schritt k) ist falsch, falls im vorherigen Schritt ein Block auf X lag. ???
% 19 :- mv(X,Y,k), on(Z,Y,k-1), bl(X;Y;Z).
%% move von X nach Y (im Schritt k) ist falsch, falls im vorherigen Schritt ein Block auf Y lag
% 20
% 21 #volatile k.
% 22
% 23 :- not on(X+1,X,k), bl(X), X < n.
%% solange keine Antwort, bis n auf n-1 ... 2 auf 1 liegt.
% 24
% 25 #base.

>>>>

 1{move(X,Y,k) : block(X) : location(Y) : X != Y}1.
\end{verbatim}
A move is of the form ``\texttt{move(X,Y,k)}'' where \texttt{X} is the block that is moved, \texttt{Y} the new position of the block and \texttt{k} the step when the moved is executed.
\newline
Now that we have exactly one move per step we need to define what happens if we move a block.
Understandably if we move block \texttt{X} to position \texttt{Y} in step \texttt{k}, the new location of the block shall be \texttt{Y}.
\begin{verbatim}
 on(X,Y,k) :- move(X,Y,k).
\end{verbatim}
For all other blocks that have not moved, the position shall be the same.
In ASP a little trick is used to describe such a condition.
We first define a little helper predicate that states that a block has moved.
\begin{verbatim}
 moved(X, k) :- move(X, Y, k).
\end{verbatim}
With this helping predicate we can define that a block is on the same position in step \texttt{k}, if it was not moved.
\begin{verbatim}
 on(X,Y,k) :- on(X,Y,k-1), block(X), location(Y), not moved(X, k).
\end{verbatim}

Now we have to restrict our moves, because only one block can be moved at the same time and it can only be moved if no other block is on top of it.
First we introduce a predicate called ``\texttt{blocked}'' for each block with another block on top of it.
\begin{verbatim}
 blocked(Y, k) :- block(Y), on(X, Y, k - 1).
\end{verbatim}
Now we can restrict the moves with an integrity constraint.
\begin{verbatim}
 :- move(X, Y, k), blocked(X, k).
\end{verbatim}
So it is not allowed to move block \texttt{X} in step k, if it is blocked.
Also you can not move a block to a position where another block is on, so only one block on top of another is allowed.
\begin{verbatim}
 :- move(X, Y, k), blocked(Y, k).
\end{verbatim}
The description of the problem is nearly finished. All that has to be done is to formulate a query.
As a constraint to our solution we want to stop if all blocks are on top of each other in reversed order.
This query shall only hold in the last step, therefore we write it in the volatile part.
As always for constraints, we use an integrity constraint to express that all solutions are false where there is not block $n+1$ on the block $n$.
\begin{verbatim}
 #volatile k.

 :- not on(X+1,X,k), block(X), X < n.
\end{verbatim}
To run this example, iclingo has to be used and the constant $n$ has to be set via command line.
A potential command line call could look like
\begin{verbatim}
 iclingo --const n=10 blocksworld.lp
\end{verbatim}
and the answer to the problem can be read from the answer set.
To clean up the answer the ``\texttt{show} and \texttt{hide}'' feature of gringo can be used.
If the following lines were added to the program, the answer set only consists of \texttt{move/3} predicates.
\begin{verbatim}
 #hide.
 #show move/3.
\end{verbatim}
